Dark Control:\\ A Unified Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-04-17 paper/paper.tex
supra-pass part 1

\author{Elvis Dohmatob, Guillaume Dumas, Danilo Bzdok\\
  \textbf{Please add your affiliation}
  %% INRIA, Parietal team, Saclay, France\\
  %% CEA, Neurospin, Gif-sur-Yvette, France\\
  %% firstname.lastname@inria.fr
  %% GD1 Institut Pasteur, Human Genetics and Cognitive Functions Unit, Paris, France
  %% GD2 CNRS UMR 3571 Genes, Synapses and Cognition, Institut Pasteur, Paris, France
  %% GD3 University Paris Diderot, Sorbonne Paris Cité, Paris, France
  %% GD4 Centre de Bioinformatique, Biostatistique et Biologie Intégrative (C3BI, USR 3756 Institut Pasteur et CNRS), Paris, France
}

@@

\begin{abstract}
The default mode network (DMN) is believed to subserve the human
baseline mental activity.
%
Many research streams converge to its likely role in evolutionarily adaptive
envisioning of mental scenarios to simulate the future.
Such functional roles become more likely due to
this brain network's highest energy consumption in the brain and
its intimate coupling with conscious awareness.
%
The present concept paper proposes a \textit{process model}
in trying to explain \textit{how} the DMN may actually implement continuous
assessment and prediction of the environment to guide action choices.
DMN function is recast in the mathematical terms of
control theory and reinforcement learning based on Markov decision processes.
We argue that our formal account of DMN function
naturally accommodates
previously proposed cognitive accounts as special cases and
parsimoniously explains
experimental findings in animals and humans.
%
A process models for the neural implemention
of continuous future predictions probably realized in the DMN
could offer insight into the statistical mechanisms underlying human behavior.
\end{abstract}

@@

\section{Introduction}
%
When left unperturbed, the human brain is not at rest.
Rather, the brain continues to metabolize large quantities of
oxygen and glucose energy to maintain
neuronal computation
in the absence of behavioral focus
\citep{kenet2003spontaneously, fiser2004small}.
The baseline energy demand is subject to surprisingly little modulations
when mentally processing environmental challenges
\citep{raichle2001pnas}. \textbf{(XXX Guillaume: What about the energy consumption compare to sleep? Danilo: I'd love to cite that
but I am not aware of such studies)}
This dark matter of brain physiology begs the question
of underlying biological purpose
\citep{raichle2006brain}.
What has early been described as the "stream of consciousness"
in psychology \citep{james1890principles}
found a potential neurobiological manifestation
in the so-called "default mode network" (DMN)
\citep{shul1997, raichle2001pnas}.
This set of some of the highest regions in
the human association cortex \citep{mesulam1998sensation, margulies2016situating}
consistently
increased in neural activity during
unfocused everyday mind wandering \citep{raichle2001pnas}.
The DMN may exert higher-order control on human behavior
following the notion of "a hierarchy of brain systems with
the DMN at the top and the salience and dorsal attention systems
at intermediate levels, above thalamic and unimodal sensory
cortex" \citep{carhart2010default}.
In the beginning of the 21st century,
brain imaging may have been the first technique
to allow for the discovery of a unique brain network that subserves
baseline mental activities
\citep{raichle2001pnas, bzdok2015resting}.

\subsection{Default mode network: higher-order control of the organism}
The human DMN accounts for a large part of the brain's baseline activity
and typically decreases in activity during psychological experiments.
The standard mode of
neural information maintenance and manipulation has
been argued to mediate evolutionarily conserved functions
\citep{brown1914nature, binder1999conceptual, buzsaki2006rhythms}.
This is all the more likely because the DMN contains hotspots
of highest energy consumption in the entire central nervous system
\citep{raichle2001pnas}.
DMN activity also persists to a substantial degree during
sleep and under anesthesia \citep{randy2008}.
Today, many authors believe that the DMN implements
probabilistic estimation of past, hypothetical, and
future events
\citep{fox2005, hassabis2007patients, schacter2007remembering, binder2009,
randy2008}.
This brain network
might have emerged to continuously predict environmental events using
mental imagery as an evolutionary advantage.
%
However, information processing in the DMN has also repeatedly
been shown to directly impact human behavior. Goal-directed task performance
improved with decreased activity in default mode areas \citep{weiss2006}
and increased DMN activity was linked to more task-independent,
yet sometimes useful thoughts
\citep{mason2007, seli2016mind}.
%
Gaining insight into the overarching DMN function is challenging because
this network appears to
simultaneously control perception-action cycles and
maintain baseline contemplations
across time, space, and content domains.

In this work,
we consider an agent faced with the choice of the next action
and being guided by rich
reenactment of really-happened, hypothetically conceived, and
predicted future events to perfect ongoing behavioral performance.
A particularly attractive computational framework
to describe, quantify, and predict autonomously acting systems like the brain
is proposed to be the combination of control theory and
reinforcement learning (RL).
It is known that, the more the external world is predictable,
the more mental activity becomes detached from the actual sensory environment
\citep{antrobus1966studies, pope1978regulation}.
Conversely,
the more the ongoing executed task is unknown and unpracticed,
the less stimulus-independent thoughts occur
\citep{filler1973daydreaming, teasdale1995stimulus, christoff2016mind}.
These ``offline'' processes may however contribute to optimized control of the organism. \textbf{(XXX Guillaume: What about the idea of model free control being also remote of environment? - danilo: hmm, let's keeping thinking about this)}
% We formalize
% a policy matrix to capture the space of possible actions that can be performed
% on the environment given the current state. A value matrix
% maps expected rewards to environmental objects and events.
% Switching between states reduces to a sequential processing model.
% In a model-based reinforcement learning (RL) approach,
% we try to understand the environment always better
% starting, at birth, with an environment that is completely unknown.
Informed by outcomes of performed action,
the DMN dynamics are constantly adapted in feedback loops
shaped by prediction error.
A DMN framework incorporating RL can naturally embed human behavior
into the tension between exploitative action with immediate gains and
exploratory action with longer-term reward schedules.

\subsection{Towards a formal account of DMN function}
We argue that the implication of the DMN in a
diversity of humans' most advanced cognitive processes
can be parsimoniously recast as prediction error minimization
based on pervasive probabilistic simulations,
thus maximizing reward outcome across multiple time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo sampling
\citep{tenenbaum2011grow}.
In this context,
\textit{control} refers to the influences that an agent exerts when interacting
with the environment to encourage preferred states.
Even (necessarily imperfect) memories, random day-time mind-wandering, and dreams during sleep
may provide meaningful building blocks to iteratively improve
the predictive DMN machinery to optimize the behavior of the organism.
%
In short, it has been proposed earlier that
the human brain's energy budget is largely dedicated to
``the development and maintenance of [a]
probabilistic model of anticipated events''
\citep{raichle2005intrinsic}.
This idea has been invigorated by
empirical evidence from
neuroscientific experiments \citep{kording2004bayesian, fiser2004small}.
The present paper proposes a
process model that satisfies this contention.

\section{Known neurobiological properties of the default mode network}
We begin by a neurobiological deconstruction of the DMN
based on experimental findings in the neuroscience literature.
This walkthrough across network nodes will
outline their individual functional profiles,
paving the way for their functional interpretation within our computational account
described in the next section.

\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_overview_DMN.pdf}
  \caption{Neurobiological overview of the DMN ... \textbf{Please add a figure caption}}
  \label{fig:neurobiological_overview_DMN}
\end{figure}
%
\subsection{The posteromedial cortex: global monitoring and information integration}
The midline structures of the DMN,
including the posteromedial cortex (PMC) and
the medial prefrontal cortex (mPFC),
are responsible for the highest turn-over
of glucose energy consumption \citep{raichle2001pnas}.
These metabolic characteristics go hand-in-hand with
brain-imaging analyses \citep{andrews2010} in
suggesting that the PMC and mPFC
potentially represent the functional backbone of the DMN.

The PMC myelinates relatively late during postnatal development in monkeys
\citep{goldman1987development}, which is generally considered to
be a sign of evolutionary sophistication \citep{flechsig1920}.
Normal and disturbed metabolic fluctuations in the
human PMC have been repeatedly related to
phenomena of changed conscious awareness \citep{cavanna2006precuneus},
including anesthesia, sleep, and coma.
%
Indeed,
the PMC has long been speculated to reflect constant computation of
environmental statistics and its internal representation
as an inner "mind's eye" \citep{cavanna2006precuneus, leech_pcc2014}.
For instance, B\'alint's syndrome is a neurological disorder of conscious
awareness that results from damage in the parietal cortex
\citep{balint1909seelenlahmung}.
Patients are plagued by an
the inability to bind various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as inability to direct action towards
currently unattended environmental objects
(i.e., optic ataxia).
This can be viewed as a high-level impairment in gathering
information about alternative objects (i.e., exploration) as well as
leveraging these environmental opportunities (i.e., exploitation).
Congruently,
the human PMC was coupled in two functional connectivity analyses
with the amygdala, involved in significance evaluation, and
the nucleus accumbens (NAc), involved in reward evaluation.
Specifically, among all parts of the PMC,
the ventral posterior cingulate cortex was
most connected to the laterobasal
nuclei group of the amygdala
\citep{bzdok2015subspecialization}.
This amygdalar subregion has been proposed to
continuously scan environmental input
for biological significance assessment.

@@

\subsection{The prefrontal cortex: stimulus-value association and action consideration}
Analogous to the PMC,
the dorsomedial PFC (dmPFC) of the DMN is believed to subserve
supra-sensory processes
across time, space, and content domains to
exert top-down control.
In comparison to the PMC, however,
the dmPFC may be closer to a
``mental sketchpad'' \citep{goldman1996prefrontal}, as it
potentially subserves the de-novo generation and binding
of meaning representations instructed by stored semantics and memories.

@@

RL is a problem-solving technique in which, through interactions with an external environment,
an agent learns to reach goals, optimize reward signals, and minimize costs
in an iterative trial-by-error fashion.
At a given moment, each taken \textit{action} $a$ triggers a change in the
\textit{state} of the environment
$s \rightarrow s'$, accompanied by environmental feedback signals as \textit{reward}
$r = r(s, a)$ (or \textit{punishment},
since it can be negative) collected by the
agent.
In this view, the environment is partially controlled by
the action of the agent and the reward can be thought
of as satisfaction -- or punishment -- accompanying the execution of
an action.
Action may be delayed to achieve substantial increases in expected reward
that can grow with time.
The environment is generally taken as stochastic,
that is, changing in random ways. In addition,
it is only \textit{partially observed}
in the sense that only part of the current state is observed by
the agent \citep{starkweather2017dopamine}.
The volatility of the environment
is realistic in a computational model which sets out
to explain DMN function.
% XXX: the following passage is already in the introduction
% Indeed, it is known that the more the external world is predictable,
% the more mental activity becomes detached from the actual sensory inputs
% \citep{antrobus1966studies, pope1978regulation},
% the more stimulus-independent thoughts occur, and
% the higher the neural activity in the DMN.
% Conversely, the more the currently executed task is unknown and unpracticed,
% the lower the DMN activity
% \citep{filler1973daydreaming, teasdale1995stimulus}.
% %
% These ``offline'' processes may however contribute to
% optimizing control of the organism.
% Such a framework of DMN funtion based on reinforcement learning
% can naturally embed human behavior
% in the tension between exploitative action with immediate gains and
% explorative action with longer-term reward schedules
% \citep{dayan2008decision}.
We argue that a functional account of the DMN based on RL
can naturally embed human behavior
in the tension between exploitative action with immediate gains and
explorative action with longer-term reward schedules
\citep{dayan2008decision}.
DMN implication in a diversity of
particularly advanced cognitive processes
can be parsimoniously explained as probabilistic simulations in form of mental scenes coupled with prediction error minimization
to calibrate action portfolios and
maximize reward outcome at different time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling
\citep{tenenbaum2011grow}.

@@

\subsection{Semantic hypothesis}
Another frequently proposed cognitive account of DMN function revolves
around forming logical associations and analogies between
the current experience and
the conceptual knowledge derived from past experiences
\citep{bar2007proactive, binder1999conceptual}.
Analogies might naturally tie incoming novel sensory stimuli to
explicit world knowledge (i.e., semantics) extracted from the environment
\citep{bar2009proactive}.
The encoding of complex environments could thus be facilitated
by association to similar states.
%
Going beyond human language itself,
semantic building blocks may provide the basis for
mentally envisioning non-existent scenarios
that would optimize behavior in the environment
by constantly simulating possible future events.
Such cognitive processes can afford
the internal generation of necessary information
that is not presented in the surrounding environment
by recombining building blocks of
concept knowledge and episodic memories
\citep{hassabis2009construction}.
Indeed, in aging humans, remembering the past and imaging the future
equally decreased in the level of detail and were associated with
concurrent deficits in forming and integrating relationships between
items \citep{addis2008age, spreng2006temporal}.
Further,
a constructive account can explain the reciprocal relationship
between an egocentric first person perspective and
an allocentric bird’s eye perspective immersed in
self-reflection, semantic associations, and autobiographical memories.
%
Cognitive aspects of egocentric-allocentric switching
are also closely related to episodic memory, language, problem solving,
planning, estimating other people's thoughts, and spatial navigation.
These neural processes are likely to
build on abstract world knowledge and logical assocations
for binding the constituent elements in mental scene construction
\citep{schacter2007remembering}.
Such scene generation processes could contribute to interpreting the
present and foretelling the future.
Mental scene construction has been proposed
to imply a distinction between
engagement in the sensory environment
and internally generated mind-wandering
\citep{buckner2007self}.
These authors stated that
``A computational model [...] will probably require a form of
regulation by which perception of the current world is suppressed
while simulation of possible alternatives are constructed,
followed by a return to perception of the present.''.
The semantic hypothesis is for instance supported by evidence in animals that
could learn a \textit{cognitive map of the environment},
even without reward incentives, and exploit it later
for other means \citep{tolman1948cognitive}.

>>>>

\author{Elvis Dohmatob, Guillaume Dumas, Danilo Bzdok\\
  INRIA, Parietal team, Saclay, France\\
  CEA, Neurospin, Gif-sur-Yvette, France\\
  Institut Pasteur, Human Genetics and Cognitive Functions Unit, Paris, France\\
  CNRS UMR 3571 Genes, Synapses and Cognition, Institut Pasteur, Paris, France\\
  University Paris Diderot, Sorbonne Paris Cité, Paris, France\\
  Centre de Bioinformatique, Biostatistique et Biologie Intégrative, Paris, France
}

@@

\begin{abstract}
The default mode network (DMN) is believed to subserve the human
baseline mental activity.
%
Many research streams converge to its likely role in evolutionarily adaptive
envisioning of mental scenarios to simulate the future.
Such functional roles become more likely due to
this brain network's highest energy consumption in the brain and
its intimate coupling with conscious awareness.
%
The present concept paper proposes a \textit{process model}
in trying to explain \textit{how} the DMN may actually implement continuous
assessment and prediction of the environment to guide action choices.
DMN function is recast in the mathematical terms of
control theory and reinforcement learning based on Markov decision processes.
We argue that our formal account of DMN function
naturally accommodates
previously proposed cognitive accounts as special cases and
parsimoniously explains
experimental findings in animals and humans.
%
A process model for the neural implemention
of continuous future predictions potentially realized in the DMN
could offer insight into the statistical mechanisms underlying human behavior.
\end{abstract}

@@

\section{Introduction}
%
When left unperturbed, the human brain is not at rest.
Rather, the brain continues to metabolize large quantities of
oxygen and glucose energy to maintain
neuronal computation
in the absence of behavioral focus
\citep{kenet2003spontaneously, fiser2004small}.
The baseline energy demand is subject to surprisingly little modulations
when mentally processing environmental challenges
\citep{raichle2001pnas}. \textbf{(XXX Guillaume: What about the energy consumption compare to sleep? Danilo: I'd love to cite that
but I am not aware of such studies)}
This dark matter of brain physiology begs the question
of underlying biological purpose
\citep{raichle2006brain}.
What has early been described as the "stream of consciousness"
in psychology \citep{james1890principles}
found a potential neurobiological manifestation
in the so-called "default mode network" (DMN)
\citep{shul1997, raichle2001pnas}.
This set of some of the highest regions in
the human association cortex \citep{mesulam1998sensation, margulies2016situating}
consistently
increase in neural activity during
unfocused everyday mind wandering \citep{raichle2001pnas}.
The DMN may subserve important functions in
higher-order control of human behavior
following the notion of "a hierarchy of brain systems with
the DMN at the top and the salience and dorsal attention systems
at intermediate levels, above thalamic and unimodal sensory
cortex" \citep{carhart2010default}.
In the beginning of the 21st century,
brain imaging may have been the first technique
to allow for the discovery of a unique brain network that subserves
baseline mental activities
\citep{raichle2001pnas, bzdok2015resting}.

\subsection{Default mode network: higher-order control of the organism}
The human DMN is responsible for a large part of the brain's baseline activity,
which typically decreases during psychological experiments.
The standard mode of
neural information maintenance and manipulation has
been argued to mediate evolutionarily conserved functions
\citep{brown1914nature, binder1999conceptual, buzsaki2006rhythms}.
In line with this, the DMN contains hotspots
of highest energy consumption in the entire central nervous system
\citep{raichle2001pnas},
which persists to a substantial degree during
sleep and under anesthesia \citep{randy2008}.
Today, many authors believe that the DMN implements
probabilistic estimation of past, hypothetical, and
future events
\citep{fox2005, hassabis2007patients, schacter2007remembering, binder2009,
randy2008}.
This brain network
might have emerged to continuously predict environmental events using
mental imagery as an evolutionary advantage.
%
However, information processing in the DMN has also repeatedly
been shown to directly impact human behavior. Goal-directed task performance
improved with decreased activity in default mode areas \citep{weiss2006}
and increased DMN activity was linked to more task-independent,
yet sometimes useful thoughts
\citep{mason2007, seli2016mind}.
%
Gaining insight into the function of the DMN is
particularly challenging because
this network appears to
simultaneously control perception-action cycles and
maintain baseline contemplations
across time, space, and content domains.

The present work takes the perspective of
a human agent faced with the choice of the next action
and being guided by rich
reenactment of really-happened, hypothetically conceived, and
estimated futures to optmize behavioral performance.
A particularly attractive computational framework
to describe, quantify, and predict autonomously acting systems like the brain
is proposed to be the combination of control theory and
reinforcement learning (RL).
It is known that, the more the external world is predictable,
the more mental activity becomes detached from the actual sensory environment
\citep{antrobus1966studies, pope1978regulation}.
Conversely,
the more the ongoing executed task is unknown and unpracticed,
the less stimulus-independent thoughts occur
\citep{filler1973daydreaming, teasdale1995stimulus, christoff2016mind}.
These ``offline'' processes may however contribute to optimized control of the organism. \textbf{(XXX Guillaume: What about the idea of model free control being also remote of environment? - danilo: hmm, let's keeping thinking about this)}
% We formalize
% a policy matrix to capture the space of possible actions that can be performed
% on the environment given the current state. A value matrix
% maps expected rewards to environmental objects and events.
% Switching between states reduces to a sequential processing model.
% In a model-based reinforcement learning (RL) approach,
% we try to understand the environment always better
% starting, at birth, with an environment that is completely unknown.
Informed by outcomes of performed action,
the DMN dynamics are constantly adapted in feedback loops
shaped by prediction error.
A DMN framework incorporating RL can naturally embed human behavior
into the tension between exploitative action with immediate gains and
exploratory action with longer-term reward schedules.

\subsection{Towards a formal account of DMN function}
We argue that DMN implication in a
diversity of humans' most advanced cognitive processes
can be parsimoniously recast as prediction error minimization
based on pervasive probabilistic simulations,
thus maximizing action outcome across multiple time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of Markov Chain Monte Carlo (MCMC) sampling
\citep{tenenbaum2011grow}.
In this context,
\textit{control} refers to the influence that an agent exerts when interacting
with the environment to encourage preferred states.
Even (necessarily imperfect) memory retrieval,
random day-time mind-wandering, and dreams during sleep
may provide meaningful building blocks to iteratively improve
the predictive DMN machinery to optimize the behavior of the organism.
%
It has indeed been proposed earlier that
the human brain's energy budget is largely dedicated to
``the development and maintenance of [a]
probabilistic model of anticipated events''
\citep{raichle2005intrinsic}.
This idea has been invigorated by
empirical evidence from
neuroscientific experiments \citep{kording2004bayesian, fiser2004small}.
The present paper proposes a
process model that satisfies this contention.

\section{Known neurobiological properties of the default mode network}
We begin by a neurobiological deconstruction of the DMN
based on experimental findings in the neuroscience literature.
This walkthrough across network nodes will
outline the individual functional profiles,
paving the way for their functional interpretation
within our formal account (next section).

\begin{figure}[!h]
  \includegraphics[width=.9\linewidth]{neurobiological_overview_DMN.pdf}
  \caption{Neurobiological overview of the DMN ... \textbf{Please add a figure caption}}
  \label{fig:neurobiological_overview_DMN}
\end{figure}
%
\subsection{The posteromedial cortex: global monitoring and information integration}
The midline structures of the DMN,
including the posteromedial cortex (PMC) and
the medial prefrontal cortex (mPFC),
are probably responsible for the highest turn-over
of glucose energy consumption \citep{raichle2001pnas}.
These metabolic characteristics go hand-in-hand with
functional neuroimaging analyses \citep{andrews2010} in
suggesting that the PMC and mPFC
potentially represent the functional backbone of the DMN.

The PMC myelinates relatively late during postnatal development in monkeys
\citep{goldman1987development}, which is generally considered to
be a sign of evolutionary sophistication \citep{flechsig1920}.
Normal and disturbed metabolic fluctuations in the
human PMC have been repeatedly related to
phenomena of changed conscious awareness \citep{cavanna2006precuneus}.
%
Indeed,
the PMC has long been speculated to reflect constant computation of
environmental statistics and its internal representation
as an inner "mind's eye" \citep{cavanna2006precuneus, leech_pcc2014}.
For instance, B\'alint's syndrome is a neurological disorder of conscious
awareness that results from damage in the parietal cortex
\citep{balint1909seelenlahmung}.
Patients are plagued by an
inability to combine various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as an inability to direct action towards
currently unattended environmental objects
(i.e., optic ataxia).
This can be viewed as a high-level impairment in gathering
information about alternative objects (i.e., exploration) as well as
leveraging these environmental opportunities (i.e., exploitation).
Congruently,
the human PMC was coupled in two functional connectivity analyses
with the amygdala, involved in significance evaluation, and
the nucleus accumbens (NAc), involved in reward evaluation.
Specifically, among all parts of the PMC,
the ventral posterior cingulate cortex was
most connected to the laterobasal
nuclei group of the amygdala
\citep{bzdok2015subspecialization}.
This amygdalar subregion has been proposed to
continuously scan environmental input
for biological significance assessment.

@@

\subsection{The prefrontal cortex: stimulus-value association and action consideration}
Analogous to the PMC,
the dorsomedial PFC (dmPFC) of the DMN is believed to subserve
supra-sensory processes
across time, space, and content domains to
exert top-down control.
In comparison to the PMC, however,
the dmPFC may be closer to a
``mental sketchpad'' \citep{goldman1996prefrontal}, as it
potentially subserves the de-novo generation and manipulation
of meaning representations instructed by stored semantics and memories.

@@

RL is a problem-solving technique in which, through interactions with an external environment,
an agent learns to reach goals, optimize reward signals, and minimize costs
in an iterative trial-by-error fashion.
At a given moment, each taken \textit{action} $a$ triggers a change in the
\textit{state} of the environment
$s \rightarrow s'$, accompanied by environmental feedback signals as \textit{reward}
$r = r(s, a)$ (or \textit{punishment},
since it can be negative) collected by the
agent.
In this view, the environment is partially controlled by
the action of the agent and the reward can be thought
of as satisfaction -- or punishment -- accompanying the execution of
an action.
Action may be delayed to achieve substantial increases in expected reward
that can grow with time.
The environment is generally taken as stochastic,
that is, changing in random ways. In addition,
it is only \textit{partially observed}
in the sense that only part of the current state is observed by
the agent \citep{starkweather2017dopamine}.
The volatility of the environment
is realistic in a computational model which sets out
to explain DMN function.
% XXX: the following passage is already in the introduction
% Indeed, it is known that the more the external world is predictable,
% the more mental activity becomes detached from the actual sensory inputs
% \citep{antrobus1966studies, pope1978regulation},
% the more stimulus-independent thoughts occur, and
% the higher the neural activity in the DMN.
% Conversely, the more the currently executed task is unknown and unpracticed,
% the lower the DMN activity
% \citep{filler1973daydreaming, teasdale1995stimulus}.
% %
% These ``offline'' processes may however contribute to
% optimizing control of the organism.
% Such a framework of DMN funtion based on reinforcement learning
% can naturally embed human behavior
% in the tension between exploitative action with immediate gains and
% explorative action with longer-term reward schedules
% \citep{dayan2008decision}.
We argue that a functional account of the DMN based on RL
can naturally embed human behavior
in the tension between exploitative action with immediate gains and
explorative action with longer-term reward schedules
\citep{dayan2008decision}.
DMN implication in a diversity of
particularly advanced cognitive processes
can be parsimoniously explained as probabilistic simulations in form of mental scenes coupled with prediction error minimization
to calibrate action portfolios and
maximize reward outcome at different time scales.
Such a purposeful optimization objective
may be solved by a stochastic approximation
based on a brain implementation of MCMC sampling
\citep{tenenbaum2011grow}.

@@

\subsection{Semantic hypothesis}
Another frequently proposed cognitive account of DMN function revolves
around forming logical associations and analogies between
the current experience and
the conceptual knowledge derived from past experiences
\citep{bar2007proactive, binder1999conceptual}.
Analogies might naturally tie incoming novel sensory stimuli to
explicit world knowledge (i.e., semantics) extracted from the environment
\citep{bar2009proactive}.
The encoding of complex environments could thus be facilitated
by association to similar states.
%
Going beyond human language itself,
semantic building blocks may provide the basis for
mentally envisioning non-existent scenarios
that would optimize behavior in the environment
by constantly simulating possible future events.
Such cognitive processes can afford
the internal generation of necessary information
that is not presented in the surrounding environment
by recombining building blocks of
concept knowledge and episodic memories
\citep{hassabis2009construction}.
Indeed, in aging humans, remembering the past and imaging the future
equally decreased in the level of detail and were associated with
concurrent deficits in forming and integrating relationships between
items \citep{addis2008age, spreng2006temporal}.
Further,
a constructive account can explain the reciprocal relationship
between an egocentric first person perspective and
an allocentric bird’s eye perspective immersed in
self-reflection, semantic associations, and autobiographical memories.
%
Cognitive aspects of egocentric-allocentric switching
are also closely related to episodic memory, language, problem solving,
planning, estimating other people's thoughts, and spatial navigation.
These neural processes are likely to
build on abstract world knowledge and logical assocations
for integrating the constituent elements in mental scene construction
\citep{schacter2007remembering}.
Such scene generation processes could contribute to interpreting the
present and foretelling the future.
Mental scene construction has been proposed
to imply a distinction between
engagement in the sensory environment
and internally generated mind-wandering
\citep{buckner2007self}.
These authors stated that
``A computational model [...] will probably require a form of
regulation by which perception of the current world is suppressed
while simulation of possible alternatives are constructed,
followed by a return to perception of the present.''.
The semantic hypothesis is for instance supported by evidence in animals that
could learn a \textit{cognitive map of the environment},
even without reward incentives, and exploit it later
for other means \citep{tolman1948cognitive}.
